\chapter{Maths and Algorithms}
\label{app:mathsandalgs}
\section{Introduction}
This appendix takes the form of a brief and essential
set of notes on various algorithms, recurrence relations, and
root finding methods, along with a discussion of important
theorems relating to continued fractions and orthogonal polynomials. 

These concepts underlie the methods discussed in the main text. 
These are mainly mathematical results
that may not be so familiar to a materials scientist.

Two monographs are the principle sources for the discussion
of orthogonal polynomials and continued fractions in this appendix. 
The first is Gabor's Szeg\"o's textbook,
Orthogonal Polynomials, which has been a standard reference
for some time. Definitions and theorems of immediate relevance 
from that work are included here. For the extended proofs and discussions 
that work should be consulted. Szeg\"o's discussion relies primarily
on the principle of orthogonality of functions on an interval, the distribution
of zeros of the functions, quadrature, and the connection with different
classes of differential equations. He does not go too 
deeply into the connection with continued fractions.

To complement Szeg\"o with an historical overview, as well as
a rigorous analysis, of the deep
connections that exist between continued fractions, 
special functions and orthogonal polynomials, Sergey 
Krushchev's work Orthogonal Polynomials and Continued Fractions: From Euler's Point
of View should be consulted \cite{khrushchev10}. 

Two tables of particular significance are included in this
appendix. The first tabulates 
the defining properties of certain classes of orthogonal polynomials:
the intervals, the weight functions, and the recurrences that can generate
them.

The second tables contains the asymptotic properties 
of the recursion coefficients due to different spectral
features: e.g. Van Hove singularities, band edges, and  band gaps. 
This table follows that described by Alphonse Magnus and we 
have extended it here with other references
from the literature.

As Haydock observed in Vol.~35,``To understand a 
particular physical system, we need only find the chain model appropriate to it."
The notion of understanding in this context means the mapping of any physical
situation to a chain model, the subsequent study of the recursion coefficents 
that are generated, determining the underlying groups that generate their behaviour, 
and establish the fixed points of the system,i.e., the singularities, band edges, 
and asymptotics all in an optimally condensed form.

\section{Euclid's Algorithm}
As pointed out by Haydock in Ref.~\cite{weaire85} the 
use of three term recurrence formulae goes back to 
ancient Greece. 

To find the greatest common divisor of two numbers $a,b$ with
$a>b$, a recurrence relation can be set up. The starting orbitals are now 
selected from ${u_{n}}\in\mathbb{N}$ a subset of the 
natural numbers (the positive integers) with ${u_1>u_2,...,1}$.

\begin{algorithm}
\caption{Euclid$'$s Algorithm}\label{alg:euclid}
\begin{algorithmic}[1]
\Procedure{Euclid}{$a,b$}\Comment{Find greatest common divisor of a and b}
\While{($u_{n-1}>u_{n}$)}
\State $q = 0$
\While{($(u_{n-1}-qu_n) > 0$)}\Comment{This is the $H\psi$ step.}
\State $u_{n+1} = -(qu_{n} - u_{n-1})$
\State $q \gets q+1$
\EndWhile\label{quotientwhile}
\State $u_{n-1} \gets u_{n}$
\State $u_{n} \gets u_{n+1}$
\EndWhile\label{euclidendwhile}
\State \textbf{return} $b$\Comment{The gcd is $u_{n}$}
\EndProcedure
\end{algorithmic}
\end{algorithm}

The use of Euclid's algorithm in the determination of 
the continued fraction representation of a real number
provides the most direct connection with the Haydock recursion
method. Instead of a continued fraction approximation to a real
number the materials scientist seeks a continued fraction
approximation to the Green's function G(E) evaluated 
at some energy.

If a number is written:
\begin{equation}
a=bq_{0} + r_{0}
\end{equation}

\begin{equation}
\frac{a}{b} = q_{0} + \frac{r_{0}}{b}
\frac{b}{r_{0}} = \frac{1}{\frac{a}{b} - q_{0}}
\end{equation}

Again the operation or `work' at each stage of the algorithm
is finding the value of the quotient $q_i$, The process
is again iterated until it terminates (in the case of a rational number)
or up to some desired precision if it is an irrational number.

%In fields of numbers euclids algorithm is used to determine
%the multiplicative inverse
%Euclid's Algorithm can be generalized 
%by analogy of Euclidean division from natural numbers
%to polynomial equations.
%\footnote{The connection between Lanczos,
%Euclidean Algorithms has been explored in pure number 
%theoretical and cryptographic literature \cite{compmath97}.
%perturbation theory of polynomials given in \cite{damanik10}.

\section{Newton's Method for Finding Roots}
\label{sec:newtonsmethod}
Newton's method is a useful way to find the roots of a polynomial
equation. If $f(\gamma_i)$ is an approximate root of an equation,
let $f(\gamma_i+h)$ be the exact root then:
%
\begin{equation}
f(\gamma_i+h)=0
\end{equation}
%
if h is small then we can expand to first order in a 
Taylor series and write:
\begin{equation}
f(\gamma_i) + h f'(\gamma_{i})= 0
\end{equation}
%
or, equivalently:
%
\begin{equation}
h = -\frac{f(\gamma_{i})}{f'(\gamma_i)}.
\end{equation}

Finally, we can improve the guess at the approximate root:
%
\begin{equation}
\gamma_{i+1} =  \gamma_{i} + h,
\end{equation}
%
and iterate the whole procedure to whatever accuracy we desire.

\section{Euler-Wallis Formula}

\begin{align}
P_{-1} = 1, & P_{0}= b_{0} \\
Q_{-1} = 1, & Q_{0}= 1 
\end{align}

\section{Sturm's Theorem}
The Sturm sequence provides a means of determining the number of 
roots for a given polynomial equation contained in a real interval.

\section{Orthogonal Polynomials}

\section{Chebyshev Polynomials}

For computational libraries implementing the procedures
for interpolating functions and the solutions of differential-integral
equations with the use of Chebyshev polynomials see Refs.\cite{nex80,daguano83}.

\section{Continued Fractions}

\section{Singularities within the Band}
The affect of singularities within a band has 
been considered by Wheeler for the case of modified moments in
Ref.~\cite{wheeler74} and in the context of the Recursion method by Ref.~\cite{hodges77}.
A formal mathematical treatment of the problem has been given in Ref.~\cite{vanlessen03} which
agrees with many of the empirical observations made by and yields explicit formula
for the recursion coefficients for weights with isolated singularities.


\section{Continued Fractions and Orthogonal Polynomials for Multiple Bands}


\section{Algorithm for Fitting a Rational Interpolation}
\label{app:ratint}
An algorithm to compute rational interpolants 
using continued fractions algorithm is described in Ref.~\cite{vidbergserene77}.
A generic function $C_N(z)$ is written as a continued fraction:
%
\begin{equation}
C_{N}(z) = \frac{a_{1}}{1+}\frac{a_{2}(z-z_{1})}{1+}...\frac{a_{N}(z-z_{N-1})}{1},
\end{equation}
%
where $z$ is the argument of the interpolating function
at the desired point, $z_{i}$ are the points the original function is sampled at,
and $a_{i}$ are the coefficients of the interpolating polynomial.
%
\begin{equation}
C_{N}(z_{i}) = u_{i}, \quad i=1,...N.
\end{equation}
%
the coefficients $a_{i}$ can be generated from the recursion relations:
%
\begin{equation}
a_{i} = g_{i}(z_{i}),\quad g_{i}(z_{i})=u_{i},\quad i=1,...,N.
\end{equation}
%
\begin{equation}
g_{p}(z) = \frac{g_{p-1}(z_{p-1}) - g_{p-1}(z)}{(z - z_{p-1})g_{p-1}(z)}, \quad p \geq 2.
\end{equation}
%
The value of the function can then be generated at the point $z$ using the relations:
%
\begin{equation}
C_{N}(z) = \frac{A_{N}(z)}{B_{N}(z)},
\end{equation}
%
where:
%
\begin{eqnarray}
A_{0}=0, \quad A_{1} = a_{1}, \quad B_{0}=B_{1}=1, \nonumber \\
A_{n+1}(z) = A_{n}(z) + (z-z_{n})a_{n+1}A_{n-1}(z), \nonumber \\
B_{n+1}(z) = B_{n}(z) + (z-z_{n})a_{n+1}B_{n-1}(z).
\end{eqnarray}

